\plaintitle

### 2016-09-21 drafts/labor/complexity.tex

It is now clear that this approach works with focused complex tasks, but
the broader wicked problem that each complex goal represents remains unsolved
\cite{rittel1973dilemmas}.
As a first example,
idea generation shows promise
\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
but there is as yet no general crowdsourced solution for
the broader goal of invention and innovation
\cite{fuge2014analysis}.
Second,
focused writing tasks are now feasible
\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,
      writingMicroTasks,agapie2015crowdsourcing}, but
there is no general solution to create
a cross-domain, high--quality crowd--powered author. 
Third,
data analysis tasks such as
clustering \cite{chilton2013cascade},
categorization \cite{andre2014crowd}, and
outlining \cite{luther2015crowdlines}
are possible, but there is no general solution for sensemaking.
It is not yet clear what insights would be required
to enable crowdsourced solutions for these broader wicked problems.
%  and
% other crowdsourced goals such as addressing climate change
% \cite{introne2011climate}.

Restricting attention to non--expert, microtask workers proved limiting.
So, \citeauthor{foundry} introduced the idea of crowdsourcing with
online paid \textit{experts} from platforms such as Upwork.
Expert crowdsourcing enables access to a much broader set of workers,
for example designers and programmers.
The same ideas can then be applied to expert ``macro-tasks''
\cite{cheng2015break,haas2015argonaut}, enabling the crowdsourcing of goals such as user--centered design \cite{foundry},
programming \cite{latoza2014microtask,Fast2016,Chen2016}, and
mentorship \cite{suzukiAtelier}.
However, there remains the open question of
how complex the work outcomes from expert crowds can be.

@@

\topic{Piecework was limited to tasks that could be clearly evaluated.}
For example, the roles required to facilitate piecework in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''~\cite{10.2307/23702539}.
\citeauthor{hart2016rise} argues that evaluation causes an ultimate complexity limit:
at some point, evaluating multidimensional work for quality
(rather than for quantity) becomes infeasible.
In his words,
``if the quality of the output is more difficult to measure than the quantity [\ldots]
then a piecework system is likely to encourage
an over--emphasis on quantity produced and an under-emphasis on quality''
\cite{hart2016rise}.
Complex work, which is often subjective to evaluate, falls victim to this criteria.

@@

\topic{The research seems to suggest that it was difficult to apply piecework to more skilled work,
particularly because maximizing the advantages of piecework seemed to reward
smaller, more constrained, more narrowly-trained tasks, and only in organizations that could pay for the equipment and people to enable it.}
For most of the \nth{19} century,
piecework was applied almost exclusively to farm and textile work.
Work was simple and widely understood
--- farm workers didn't need to be trained on how to plow fields, or birth foals;
seamstresses knew how to sew together denim
\cite{10.2307/2338394,riisOtherSideLives}.

@@

\subsubsection{\implication}
% \topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
% it also signals some of the ultimate limitations of crowd work and piecework in general.}
% While the threshold preventing task requesters from utilizing piecework
% has dropped thanks to affordances of the Internet, the ceiling on task complexity hasn't moved significantly.
% \msb{is that your prediction? I would argue against the fact that it hasn't moved significantly, crowdforge did far more complex work, as did flash teams and flash orgs}
% If we're to make use of \citeauthor{Brown01041990}'s prescriptions,
% we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes.
Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
Beyond this, further improvements would most likely come from replicating the success of narrowly-slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
To some extent, an argument can be made that
MOOCs and other online education resources
provide crowd workers with the resources that they need, but 
it remains to be seen whether that work will be appropriately valued, let alone
properly interpreted by task solicitors
\cite{aguaded2013mooc}.
If we can overcome this obstacle,
we might be able to empower more crowd workers to do complex work such as engineering and metalworking,
rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work unless it makes common expertise more broadly available.

Will the shift from human managers to Turing-complete algorithms raise the complexity ceiling? 
By the Turing test, the algorithms would be at best indistinguishable from human piecework clerks and foremen.
%At worst, as now, they are far simpler~\cite{uberAlgorithm}.
So in terms of enabling coordination, algorithmic management is unlikely to directly raise the ceiling beyond what piecework could achieve.
However, as a resource constraint, algorithms are a fixed cost and not a per-person cost like human managers.
So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to afford crowd work.
This shift may enable complex goals that were not cost-effective before to become feasible.
However, because algorithms remain far from replicating all of the foremen's responsibilities, most likely is a middle ground in which crowd work re-introduces the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters \cite{MaliciousCrowdworkersGadiraju}.
% and develop more considerate platforms as \citeauthor{takingAHITMcInnis} advocate
% \cite{takingAHITMcInnis}.
% \ali{What will happen if we don't do this?
% Maybe this is where I should bring up
% the ``market for lemons'' metaphor and say that
% the market will just continue to decline in quality until
% nobody is willing to rely on it either as a source of data/information (requesters), or
% a means of sustaining oneself financially (workers).}

>>>>

It is now clear that this approach works with focused complex tasks, but
the broader wicked problem that each complex goal represents remains unsolved
\cite{rittel1973dilemmas}.
As a first example,
idea generation shows promise
\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
but there is as yet no general crowdsourced solution for
the broader goal of invention and innovation
\cite{fuge2014analysis}.
Second,
focused writing tasks are now feasible
\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,
      writingMicroTasks,agapie2015crowdsourcing}, but
there is no general solution to create
a cross--domain, high--quality crowd--powered author. 
Third,
data analysis tasks such as
clustering \cite{chilton2013cascade},
categorization \cite{andre2014crowd}, and
outlining \cite{luther2015crowdlines}
are possible, but there is no general solution for sensemaking.
It is not yet clear what insights would be required
to enable crowdsourced solutions for these broader wicked problems.
%  and
% other crowdsourced goals such as addressing climate change
% \cite{introne2011climate}.

Restricting attention to non--expert, microtask workers proved limiting.
So, \citeauthor{foundry} introduced the idea of crowdsourcing with
online paid \textit{experts} from platforms such as Upwork.
Expert crowdsourcing enables access to a much broader set of workers,
for example designers and programmers.
The same ideas can then be applied to expert ``macro--tasks''
\cite{cheng2015break,haas2015argonaut}, enabling the crowdsourcing of goals such as user--centered design \cite{foundry},
programming \cite{latoza2014microtask,Fast2016,Chen2016}, and
mentorship \cite{suzukiAtelier}.
However, there remains the open question of
how complex the work outcomes from expert crowds can be.

@@

\topic{Piecework was limited to tasks that could be clearly evaluated.}
For example, the roles required to facilitate piecework in the early \nth{20} century included ``piecework clerks, inspectors, and `experts'''~\cite{10.2307/23702539}.
\citeauthor{hart2016rise} argues that evaluation causes an ultimate complexity limit:
at some point, evaluating multidimensional work for quality
(rather than for quantity) becomes infeasible.
In his words,
``if the quality of the output is more difficult to measure than the quantity [\ldots]
then a piecework system is likely to encourage
an over--emphasis on quantity produced and an under--emphasis on quality''
\cite{hart2016rise}.
Complex work, which is often subjective to evaluate, falls victim to this criteria.

@@

\topic{The research seems to suggest that it was difficult to apply piecework to more skilled work,
particularly because maximizing the advantages of piecework seemed to reward
smaller, more constrained, more narrowly--trained tasks, and only in organizations that could pay for the equipment and people to enable it.}
For most of the \nth{19} century,
piecework was applied almost exclusively to farm and textile work.
Work was simple and widely understood
--- farm workers didn't need to be trained on how to plow fields, or birth foals;
seamstresses knew how to sew together denim
\cite{10.2307/2338394,riisOtherSideLives}.

@@

\subsubsection{\implication}
% \topic{The piecework literature gives us a template for pushing the boundaries of complexity in piecework, but
% it also signals some of the ultimate limitations of crowd work and piecework in general.}
% While the threshold preventing task requesters from utilizing piecework
% has dropped thanks to affordances of the Internet, the ceiling on task complexity hasn't moved significantly.
% \msb{is that your prediction? I would argue against the fact that it hasn't moved significantly, crowdforge did far more complex work, as did flash teams and flash orgs}
% If we're to make use of \citeauthor{Brown01041990}'s prescriptions,
% we would benefit from finding ways to decompose varied tasks into homogeneous microtasks. \msb{isn't that what we've been doing all along?}
Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes.
Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
of human computation \cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework \cite{hart2013rise}.
To some extent, an argument can be made that
MOOCs and other online education resources
provide crowd workers with the resources that they need, but 
it remains to be seen whether that work will be appropriately valued, let alone
properly interpreted by task solicitors
\cite{aguaded2013mooc}.
If we can overcome this obstacle,
we might be able to empower more crowd workers to do complex work such as engineering and metalworking,
rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work unless it makes common expertise more broadly available.

Will the shift from human managers to Turing--complete algorithms raise the complexity ceiling? 
By the Turing test, the algorithms would be at best indistinguishable from human piecework clerks and foremen.
%At worst, as now, they are far simpler~\cite{uberAlgorithm}.
So in terms of enabling coordination, algorithmic management is unlikely to directly raise the ceiling beyond what piecework could achieve.
However, as a resource constraint, algorithms are a fixed cost and not a per--person cost like human managers.
So in terms of accessibility, algorithms will allow a broader class of organizations and individuals to afford crowd work.
This shift may enable complex goals that were not cost--effective before to become feasible.
However, because algorithms remain far from replicating all of the foremen's responsibilities, most likely is a middle ground in which crowd work re--introduces the human element to management in a more targeted way~(e.g.,~\cite{haas2015argonaut,kulkarni2012mobileworks,crowdguilds}).
This move will require resolving the tension between workers and perilously antagonistic managers, as \citeauthor{10.2307/2118435} suggest, to break a toxic cycle of mistrustful requesters \cite{MaliciousCrowdworkersGadiraju}.
% and develop more considerate platforms as \citeauthor{takingAHITMcInnis} advocate
% \cite{takingAHITMcInnis}.
% \ali{What will happen if we don't do this?
% Maybe this is where I should bring up
% the ``market for lemons'' metaphor and say that
% the market will just continue to decline in quality until
% nobody is willing to rely on it either as a source of data/information (requesters), or
% a means of sustaining oneself financially (workers).}

### 2016-09-21 drafts/labor/decomposition.tex

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
% \onlyinsubfile{
% \usepackage{xr-hyper}
% \usepackage{hyperref}
% \externaldocument{complexity}
% \externaldocument{relationships}
% % \externaldocument{decomposition}
% }
\begin{document}

@@

\topic{The general lesson has been that the more micro the task, and the more fine the decomposition, the greater the risk that workers lose context necessary to perform the work well.}
For example, workers edit adjacent paragraphs in inconsistent ways~\cite{bernsteinSoylent,Kim2017}, interpret tasks in different ways~\cite{kairam2016parting}, and exhibit lower motivation~\cite{Kinnaird:2012:WTM:2389176.2389219} without sufficient context.
Research has sought to ameliorate this issue by designing workflows help workers ``act with global understanding when each contributor only has access to local views''~\cite{verroios2014context}, typically by automatically or manually generating higher-level representations for the workers to reflect on~\cite{chilton2013cascade,verroios2014context,Kim2017}.

@@

\topic{Piecework became an important factor in the war effort for the Second World War,
cementing its role not only in American factories, but in industrial work around the world.}
The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.
As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
He attributes this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.
One might reflect on the observation that ``Rosie the Riveter''
--- an icon of mid--\nth{20} century America who represented empowerment and opportunity for women \cite{honey1985creating} ---
was herself a pieceworker
\cite{davies2014origins}.

\topic{The research community relating to
piecework and labor
has been wrestling with the decomposition of work for centuries.}
The beginnings of
systematic task decomposition
stretch back as far as the \nth{19} century,
when Airy employed young boys at the Greenwich Observatory who
``possessed the basic skills of mathematics, including
`Arithmetic, the use of Logarithms, and Elementary Algebra'~''
to compute astronomical phenomena
\cite{grier2013computers}.
%These workers became the first \textit{computers}.

The work that Airy solicited resonates with modern crowd work for several reasons.
First, work output was quickly verifiable;
Airy could assign variably skilled workers to compute values,
and have other workers check their work.
Second, tasks were discrete --- that is, independent from one another.
Finally, knowledge of the full scope of the project
--- indeed, knowledge of anything more than the problem set at hand ---
was unnecessary.

@@

\topic{Where measurement and instrumentation were limiting factors for historical piecework, computation has changed the situation so that a dream of scientific management and Taylorism
--- to measure every motion at every point throughout the workday and beyond ---
is not only doable, but trivial
\cite{waltz2012quantified}.}
Where \citeauthor{10.2307/23702539} directly implicates measurement as
% cites as
preventing scientific management from being fully utilized,
%the difficulty of tracking work \& workers, no longer exists
%\cite{10.2307/23702539}.}
modern crowd work is measuring and modeling every click, scroll, and keyboard event~\cite{rzeszotarski2011instrumenting,rzeszotarski2012crowdscape}.
The result is that on--demand work can articulate and track far more carefully than piecework historically could.

\topic{A second shift is  
%A number of factors in crowd work are different from piecework, chief among them being 
the relative ease with which the metaphorical ``assembly line'' can be changed.}
% Computers make it possible to switch from one task to another unlike any arbitrary manufacturing factory possibly could;
%modern on--demand workers on AMT and other platforms can more readily do any number of
%different \textit{types} of tasks in the span of just a few minutes
Historical manufacturing equipment could not quickly be assembled, edited, and redeployed.\msb{cite for historical piecework equipment being a large time + money investment?}
In contrast, today system--designers can share, modify, and instantiate environments
like sites of labor in a few lines of code
\cite{lessig2006code,turkitLittle}.
This opportunity has spurred an entire body of work investigating the effects of
ordering,
pacing,
interruptions, and
other factors in piecework that would have been
all but impossible to manipulate as few as 20 years ago
\cite{dai2015and,Cai:2016:CRI:2858036.2858237,cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.

Third, modern crowd work has sliced work to such small scales that the marginal activities
--- things like finding work and cognitive task switching ---
have become relatively large compared to the tasks themselves
\cite{taskSearch}.
In the historical case of piecework,
moving metallurgical tools, mining equipment, or
other industry materials would have been prohibitively difficult and slow;
workers were encouraged to specialize in a single set of tasks,
allowing pieceworkers to sequence their tasks optimally on their own
\cite{hart2013rise}.
The result is that crowd workers are more free agents than historically was the case.
However, because they spend significant time searching for tasks, the piece rate is less a good estimate of take-home earnings than before.

% Rather than fall into the trap that \citeauthor{irani2015cultural} warns of,
% --- one which where crowd workers are rendered as
% ``modular, protocol--defined computational services'' ---
% we may yield better results from crowd work if we think of workers as similar to
% specialized, repurposable tools
% \cite{irani2015cultural}.
% \ali{feeling meh about this argument\dots}

\subsubsection{\implication}
If measurement precision limited the depth of decomposition for piecework historically, as \citeauthor{10.2307/23702539} argues, then modern on--demand work stands to become far more finely-sliced and highly decomposed than ever before.
Online tools make measurement and validation so easy~\cite{rzeszotarski2011instrumenting} that these aspects of piecework are solved, or near enough that they no longer limit task decomposition.
Now, not just tasks, but entire workers' histories~\cite{hata2017glimpse}, can be collected and analyzed in detail.
% crowd work research today is on the right track to investigate pipelining and meta--task design.
% That is, investigating better work discovery methods, producing tools for workers to make more informed decisions
% \cite[see, for example,][]{turkopticon}.

However, decomposition has hit a second bottleneck: cognition. 
Task switching costs and other cognitive costs make it difficult to work tasks so far decontextualized from their original intention \cite{delayAndOrderLasecki}.
There will of course be tasks that can be decomposed without much context, and these will form the most fine-grained of microtasks.
However, other tasks cannot be freed from context --- for example, logo design requires a deep understanding of the client and their goals.
In part due to this limitation, 99designs workers often recycle old designs rather than make new ones for each client~\cite{araujo201399designs}.

So, ultimately, the levels of decomposition are likely to follow the contours of context required. Low-context work will be extremely highly decomposed. High-context work will continue to be limited.

>>>>

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
% \onlyinsubfile{
% \usepackage{xr--hyper}
% \usepackage{hyperref}
% \externaldocument{complexity}
% \externaldocument{relationships}
% % \externaldocument{decomposition}
% }
\begin{document}

@@

\topic{The general lesson has been that the more micro the task, and the more fine the decomposition, the greater the risk that workers lose context necessary to perform the work well.}
For example, workers edit adjacent paragraphs in inconsistent ways~\cite{bernsteinSoylent,Kim2017}, interpret tasks in different ways~\cite{kairam2016parting}, and exhibit lower motivation~\cite{Kinnaird:2012:WTM:2389176.2389219} without sufficient context.
Research has sought to ameliorate this issue by designing workflows help workers ``act with global understanding when each contributor only has access to local views''~\cite{verroios2014context}, typically by automatically or manually generating higher--level representations for the workers to reflect on~\cite{chilton2013cascade,verroios2014context,Kim2017}.

@@

\topic{Piecework became an important factor in the war effort for the Second World War,
cementing its role not only in American factories, but in industrial work around the world.}
The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.
As discussed earlier, \citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
He attributes this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who represented empowerment and opportunity for women \cite{honey1985creating},
was a pieceworker
\cite{davies2014origins}.

\topic{The research community relating to
piecework and labor
has been wrestling with the decomposition of work for centuries.}
The beginnings of
systematic task decomposition
stretch back as far as the \nth{19} century,
when Airy employed young boys at the Greenwich Observatory who
``possessed the basic skills of mathematics, including
`Arithmetic, the use of Logarithms, and Elementary Algebra'~''
to compute astronomical phenomena
\cite{grier2013computers}.
The work that Airy solicited resonates with modern crowd work for several reasons.
First, work output was quickly verifiable;
Airy could assign variably skilled workers to compute values,
and have other workers check their work.
Second, tasks were discrete --- that is, independent from one another.
Finally, workers could be trained on a very narrow subset of mathematical skills to be
sufficiently qualified to do this work.

@@

\topic{Where measurement and instrumentation were limiting factors for historical piecework, computation has changed the situation so that a dream of scientific management and Taylorism
--- to measure every motion at every point throughout the workday and beyond ---
is not only doable, but trivial
\cite{waltz2012quantified}.}
Where \citeauthor{10.2307/23702539} directly implicates measurement as
% cites as
preventing scientific management from being fully utilized,
%the difficulty of tracking work \& workers, no longer exists
%\cite{10.2307/23702539}.}
modern crowd work is measuring and modeling every click, scroll, and keyboard event
\cite{rzeszotarski2011instrumenting,rzeszotarski2012crowdscape}.
The result is that on--demand work can articulate and track far more carefully than piecework historically could.

\topic{A second shift is  
the relative ease with which the metaphorical ``assembly line'' can be changed.}
Historical manufacturing equipment could not quickly be assembled, edited, and redeployed
\cite{hu1961parallel}.
In contrast, today system--designers can share, modify, and instantiate environments
like sites of labor in a few lines of code
\cite{lessig2006code,turkitLittle}.
This opportunity has spurred an entire body of work investigating the effects of
ordering,
pacing,
interruptions, and
other factors in piecework that would have been
all but impossible to manipulate as few as 20 years ago
\cite{dai2015and,Cai:2016:CRI:2858036.2858237,cheng2015break,measuringCrowdsourcingCheng,embracingErrorKrishna}.

Third, modern crowd work has sliced work to such small scales that the marginal activities
--- things like finding work and cognitive task switching ---
have become large relative to the tasks themselves
\cite{taskSearch}.
In the historical case of piecework,
moving metallurgical tools, mining equipment, or
other industry materials would have been prohibitively difficult and slow;
workers were encouraged to specialize in a single set of tasks,
allowing pieceworkers to sequence their tasks optimally on their own
\cite{hart2013rise}.
The result is that crowd workers are more free agents than historically was the case.
However, because they spend significant time searching for tasks, the piece rate is less a good estimate of take--home earnings than before.

\subsubsection{\implication}
If measurement precision limited the depth of decomposition for piecework historically, as \citeauthor{10.2307/23702539} argues, then modern on--demand work stands to become far more finely--sliced and highly decomposed than ever before.
Online tools make measurement and validation so easy~\cite{rzeszotarski2011instrumenting} that these aspects of piecework are solved, or near enough that they no longer limit task decomposition.
Now, not just tasks, but entire workers' histories~\cite{hata2017glimpse}, can be collected and analyzed in detail.

However, decomposition has hit a second bottleneck: cognition. 
Task switching costs and other cognitive costs make it difficult to work tasks so far decontextualized from their original intention \cite{delayAndOrderLasecki}.
There will of course be tasks that can be decomposed without much context, and these will form the most fine--grained of microtasks.
However, other tasks cannot be freed from context --- for example, logo design requires a deep understanding of the client and their goals.
In part due to this limitation, 99designs workers often recycle old designs rather than make new ones for each client~\cite{araujo201399designs}.

So, ultimately, the levels of decomposition are likely to follow the contours of context required. Low--context work will be extremely highly decomposed. High--context work will continue to be limited.
